Exact asymptotic expressions for local clustering in hyperbolic random graphs \\ \TM{ I prefer not to use the term ``hyperbolic random graph". Better : KPKVB random graph or, in the title,``a hyperbolic model of complex networks"}

### 2019-01-19 transition.tex
Performed the splitting of cases for k and responded to/implemented several comments from Tobias.

Now that we have a complete picture of the behavior of the local clustering coefficient and function in the limit model $G_\Pcal$ we will proceed to relate this to the clustering in the hyperbolic model $G_{\H,n}$. The first step is to relate the clustering in the infinite model to that of the finite box model $G_{\Pcal, n}$. Recall that $G_{\Pcal,n}$ is obtained by restricting the Poisson Point Process $\PPP$ to the box $\Rcal_n = (-I_n, I_n] \times (0, R_n]$, with $I_n = \frac{\pi}{2} e^{R_n/2}$ (we write $\Pcal_n$ for this process), and connecting two points $p_1, p_2 \in \Rcal_n$ if $|x_1 - x_2|_{\pi e^{R_n/2}} \le e^{(y_1 + y_2)/2}$ where we identified the left and right boundary of $\Rcal_n$. 
\TM{ This last bit of the sentence is already contained in the fact we us $|.|_{\pi e^{R_n/2}}$. To me it does not add anything, except possibly confusion. }
See Section~\ref{ssec:finite_model} for more details.

In this section we shall compute the asymptotic difference between triangle counts in the finite and infinite model. First we recall the definition of $\Kcal_{C}(k_n)$
\[
	\Kcal_C(k_n) = \left\{p \in \R : \frac{k_n - C \kappa_n}{\xi_{\alpha,\nu}} \vee 0 \le e^{\frac{y}{2}}
	\le \frac{k_n + C \kappa_n}{\xi_{\alpha,\nu}} \wedge e^{R_n/2} \right\},	
\]
\TM{ Typo : $p\in\R$. Do you mean $\Rcal_n$? Or maybe 
$\R \time \R^+$? Still not sure that max and min in defn add anything. }
with
\[
	\kappa_n := \begin{cases}
		\log(n) &\mbox{if } k_n = \bigT{1},\\
		\sqrt{k_n \log(k_n)} &\mbox{else.}
	\end{cases}
\]
\TM{ Same issue as before. }
In addition we note that the function $\phi_n(y)$ in Lemma~\ref{lem:average_degree_P_n} satisfies
\[
	\phi_n(y) = \bigT{e^{-(\alpha - \frac{1}{2})(R_n - y)}},
\]
which yields the following useful corollary to Lemma~\ref{lem:average_degree_P_n}.

\begin{corollary}\label{cor:average_degree_P_n_K}
Let $\alpha > \frac{1}{2}$, $C > 0$ and $k_n$ be any sequence satisfying $k_n = \smallO{n^{\frac{1}{2\alpha - 1}}}$. \TM{ Typo? $2\alpha-1$ should be $2\alpha+1$? } 
Then, for all $p \in \Kcal_{C}(k_n)$,
\[
	\mu_{\alpha,\nu ,n}(B_{\Pcal,n}(p)) = \mu_{\alpha,\nu}(B_\Pcal(p))\left(1 - \bigT{k_n^{2\alpha - 1} n^{-(2\alpha - 1)}}\right).
\]
\TM{ I don't follow. Spell it out. }

@@

We now turn to the task of calculating the expected number of triangles of a node at height $y$, for both the infinite model and the finite box model. Let $p_0 = (0,y)$ and recall that 
\[
	\Exp{T_{\Pcal,n}(p_0)} = \frac{1}{2}\iint_{\Rcal_n^2} T_{\Pcal,n}(p_0,p_1,p_2)
	f_{\alpha,\nu}(x_1,y_1) f_{\alpha,\nu}(x_1,y_1) \, dx_1 \, dx_2 \, dy_1 \, dy_2,
\]
where
\[
	T_{\Pcal,n}(p_0,p_1,p_2) = \ind{p_1 \in B_{\Pcal,n}(p)}\ind{p_2 \in B_{\Pcal,n}(p)}\ind{p_2 \in B_{\Pcal,n}(p_1)}.
\]
\TM{ $p$ is supposed to be $p_0$ I assume. }
The difference between the indicator $\ind{p_1 \in B_{\Pcal,n}(p)}$ in the finite box model and $\ind{p_1 \in B_{\Pcal}(p)}$ is that in $G_{\Pcal,n}(\alpha,\nu)$ we identified the boundaries of the interval $[-\frac{\pi}{2}e^{R_n/2}, \frac{\pi}{2} e^{R_n/2}]$. \TM{ AND we stop at height $y=R_n$! (may be significant at some point) } 
It is clear that for any $0 \le y \le R_n$ we have that $\BallPon{p_0} = \BallPo{p_0} \cap \Rcal_n$. However, when we take another point $p^\prime \in \BallPon{p}$ then it could happen that there are points in the intersection $\BallPon{p} \cap \BallPon{p^\prime}$ that are not in $\BallPo{p} \cap \BallPo{p^\prime}$. Let us denote this region by $\mathcal{T}_{\Pcal \Delta \Pcal_n}(p,p^\prime)$. Then, any $p_2 \in \mathcal{T}_{\Pcal \Delta \Pcal_n}(p,p^\prime)$ creates a triangle with $p$ and $p^\prime$ in $G_{\Pcal,n}(\alpha,\nu)$ that is not present in $G_{\Pcal}(\alpha, \nu)$.

Define
\begin{equation}\label{eq:def_tilde_triangle_indicator}
	\widetilde{T}_{\Pcal,n}(p_0,p_1,p_2) = \ind{p_1 \in B_{\Pcal,n}(p)}\ind{p_2 \in B_{\Pcal,n}(p)}\ind{p_2 \in B_{\Pcal}(p_1) \cap \Rcal_n}
\end{equation}
\TM{ $p$ should be $p_0$ I supppose. }
Then
\[
	\sum_{p_1,p_2 \in \Pcal_n \setminus (0,y)}^{\ne} T_{\Pcal,n}(p_0,p_1,p_2) - \widetilde{T}_{\Pcal,n}(p_0,p_1,p_2)
	= \sum_{p_1,p_2 \in \Pcal_n \setminus (0,y)}^{\ne} \ind{p_1 \in \BallPon{p_0}} \ind{p_2 \in \mathcal{T}_{\Pcal \Delta \Pcal_n}(p_0,p_1)},
\]
\TM{ Note you cannot really force/assume $p_0 = (0,y)$ is in your PPP $\Pcal_n$. (The summation over $\Pcal_n \setminus (0,y)$ suggests this is what you do.) For taking the expectation of sort of thing, Palm theory / Mecke are used typically. At this point though you can again still just say we add $p_0$ to the PPP. }
where the sums are over all distinct pairs $(p_1, p_2) \in \Pcal_n \setminus \{(0,y)\}$.

@@

To further analyze this, let us introduce some notation. For any $p = (x,y) \in \Rcal_n$ we will define the left and right boundary functions as, respectfully,\TM{ respectively? }
\begin{align}
	b_p^-(z) &= \begin{cases}
		2 \log\left(x-z\right) - y &\mbox{if }  -\frac{\pi}{2} e^{R_n/2} \le z \le x - e^{y/2}  \\
		2\log\left(\pi e^{R_n/2} + x - z\right) - y 
			&\mbox{if } x - e^{(y + R_n)/2} + \pi e^{R_n/2} \le z \le \frac{\pi}{2} e^{R_n/2}\\
		0 &\mbox{else}
	\end{cases}\\
	b_p^+(z) &= \begin{cases}
		2 \log\left(z-x\right) - y &\mbox{if } x + e^{y/2} \le z \le \frac{\pi}{2} e^{R_n/2} \\
		2\log\left(\pi e^{R_n/2} + z - x\right) - y 
			&\mbox{if } -\frac{\pi}{2} e^{R_n/2} \le z \le x + e^{(y + R_n)/2} - \pi e^{R_n/2}\\
		0 &\mbox{else}
	\end{cases}
\end{align}
Note that these functions describe the boundaries of the ball $\BallPon{p}$. In particular, $p^\prime = \BL{(x^\prime, y^\prime)} \in \BallPon{p}$ if and only if $y^\prime \ge \min\left\{b_p^-(x^\prime), b_p^+(x^\prime)\right\}$.

For the remainder of the analysis we restrict ourselves to the case where $x^\prime > 0$. Due to symmetry the situation for $x^\prime < 0$ is the same. \TM{ does this not depend on $p$? } 
There are two important points in the box $\Rcal_n$. These are the intersection between the left boundary of $p^\prime$ and the right boundary of $p^\prime$, as it continues from the left side of the box, and the left boundary of $p$. We denote by $(x^\ast(p^\prime), y^\ast(p^\prime))$ the intersection between the left and right boundary of $p^\prime$ and by $(\hat{x}(p,p^\prime), \hat{y}(p,p^\prime))$ the intersection between the left boundary of $p$ and the right boundary of $p^\prime$, see Figure \ref{fig:comparing_triangles_diff_analysis}. After some simple algebra \TM{ I am not a big fan of writing "after some simple algebra" or similar. This is how errors can creep into your argument. Maybe we can just give the details? } we obtain the following expressions for the coordinates of these two points
\begin{align*}
	x^\ast(p^\prime) &= x^\prime - \frac{\pi}{2} e^{R_n/2}\\
	y^\ast(p^\prime) &= 2\log\left(\frac{\pi}{2}e^{R_n/2}\right) - y^\prime\\
	\hat{x}(p,p^\prime) &= \frac{x^\prime - \pi e^{R_n/2}}{1 + e^{(y^\prime - y)/2}} \\
	\hat{y}(p,p^\prime) &= 2 \log\left(\frac{\pi e^{R_n/2} - x^\prime}{e^{y/2} + e^{y^\prime/2}}\right)
\end{align*}

>>>>

Now that we have a complete picture of the behavior of the local clustering coefficient and function in the limit model $G_\Pcal$ we will proceed to relate this to the clustering in the hyperbolic model $G_{\H,n}$. The first step is to relate the clustering in the infinite model to that of the finite box model $G_{\Pcal, n}$. Recall that $G_{\Pcal,n}$ is obtained by restricting the Poisson Point Process $\PPP$ to the box $\Rcal_n = (-I_n, I_n] \times (0, R_n]$, with $I_n = \frac{\pi}{2} e^{R_n/2}$ (we write $\Pcal_n$ for this process), and connecting two points $p_1, p_2 \in \Rcal_n$ if $|x_1 - x_2|_{\pi e^{R_n/2}} \le e^{(y_1 + y_2)/2}$. We recall that by definition of the norm $|.|_{\pi e^{R_n/2}}$ the left and right boundaries of $\Rcal_n$ are identified. See Section~\ref{ssec:finite_model} for more details. 
\TM{ This last bit of the sentence is already contained in the fact we us $|.|_{\pi e^{R_n/2}}$. To me it does not add anything, except possibly confusion. } \PvdH{I have updated this and just added a reminder of this.}

\begin{remark}[Diverging $k_n$]
Throughout this section $\{k_n\}_{n \ge 1}$ will denote an sequence of integers satisfying $k_n \to \infty$ and $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$, as $n \to \infty$.
\end{remark}

In this section we shall compute the asymptotic difference between triangle counts in the finite and infinite model. First we recall the definition of $\Kcal_{C}(k_n)$
\[
	\Kcal_C(k_n) = \left\{p \in \Rcal : \frac{k_n - C \sqrt{k_n \log(k_n)}}{\xi_{\alpha,\nu}} \le e^{\frac{y}{2}}
	\le \frac{k_n + C \sqrt{k_n \log(k_n)}}{\xi_{\alpha,\nu}} \right\},	
\]
\TM{ Typo : $p\in\R$. Do you mean $\Rcal_n$? Or maybe 
$\R \times \R^+$? Still not sure that max and min in defn add anything. } \PvdH{It was a typo and is corrected. I have removed the max and min.}
%with
%\[
%	\kappa_n := \begin{cases}
%		\log(n) &\mbox{if } k_n = \bigT{1},\\
%		\sqrt{k_n \log(k_n)} &\mbox{else.}
%	\end{cases}
%\]
%\TM{ Same issue as before. }
In addition we note that the function $\phi_n(y)$ in Lemma~\ref{lem:average_degree_P_n} satisfies
\[
	\phi_n(y) = \bigT{e^{-(\alpha - \frac{1}{2})(R_n - y)}},
\]
as $R_n - y \to \infty$. Since $e^{y/2} = \bigT{k_n} = \smallO{e^{R_n}}$ as $n \to \infty$, uniformly on $\Kcal_C$, it follows that $\phi_n(y) = \bigT{k_n^{2\alpha - 1} n^{-(2\alpha - 1)}}$, uniformly on $\Kcal_C$. This yields the following useful corollary to Lemma~\ref{lem:average_degree_P_n}.

\begin{corollary}\label{cor:average_degree_P_n_K}
Let $\alpha > \frac{1}{2}$, $C > 0$. Then, for all $p \in \Kcal_{C}(k_n)$,
\[
	\mu_{\alpha,\nu ,n}(B_{\Pcal,n}(p)) = \mu_{\alpha,\nu}(B_\Pcal(p))\left(1 - \bigT{k_n^{2\alpha - 1} n^{-(2\alpha - 1)}}\right).
\]
\TM{ I don't follow. Spell it out. }\PvdH{I added additional explanation above. Please see if this solves the confusion.}

@@

We now turn to the task of calculating the expected number of triangles of a node at height $y$, for both the infinite model and the finite box model. Recall that we added a typical point $p_0 = (0,y)$, with exponentially distributed height, to the Poisson Point Process and that
\[
	\Exp{T_{\Pcal,n}(p_0)} = \frac{1}{2}\iint_{\Rcal_n^2} T_{\Pcal,n}(p_0,p_1,p_2)
	f_{\alpha,\nu}(x_1,y_1) f_{\alpha,\nu}(x_1,y_1) \, dx_1 \, dx_2 \, dy_1 \, dy_2,
\]
where
\[
	T_{\Pcal,n}(p_0,p_1,p_2) = \ind{p_1 \in B_{\Pcal,n}(p_0)}\ind{p_2 \in B_{\Pcal,n}(p_0)}\ind{p_2 \in B_{\Pcal,n}(p_1)}.
\]
\TM{ $p$ is supposed to be $p_0$ I assume. }\PvdH{Indeed. It has been corrected.}
The difference between the indicator $\ind{p_1 \in B_{\Pcal,n}(p)}$ in the finite box model and $\ind{p_1 \in B_{\Pcal}(p)}$ is that in $G_{\Pcal,n}(\alpha,\nu)$ we identified the boundaries of the interval $[-\frac{\pi}{2}e^{R_n/2}, \frac{\pi}{2} e^{R_n/2}]$ and we stop at height $y = R_n$. It is clear that for any $0 \le y \le R_n$ we have that $\BallPon{p_0} = \BallPo{p_0} \cap \Rcal_n$. However, when we take another point $p^\prime \in \BallPon{p}$ then it could happen that there are points in the intersection $\BallPon{p} \cap \BallPon{p^\prime}$ that are not in $\BallPo{p} \cap \BallPo{p^\prime}$. Let us denote this region by $\mathcal{T}_{\Pcal \Delta \Pcal_n}(p,p^\prime)$. Then, any $p_2 \in \mathcal{T}_{\Pcal \Delta \Pcal_n}(p,p^\prime)$ creates a triangle with $p$ and $p^\prime$ in $G_{\Pcal,n}(\alpha,\nu)$ that is not present in $G_{\Pcal}(\alpha, \nu)$.

Define
\begin{equation}\label{eq:def_tilde_triangle_indicator}
	\widetilde{T}_{\Pcal,n}(p_0,p_1,p_2) = \ind{p_1 \in B_{\Pcal,n}(p_0)}\ind{p_2 \in B_{\Pcal,n}(p_0)}\ind{p_2 \in B_{\Pcal}(p_1) \cap \Rcal_n}
\end{equation}
Then
\[
	\sum_{p_1,p_2 \in \Pcal_n}^{\ne} T_{\Pcal,n}(p_0,p_1,p_2) - \widetilde{T}_{\Pcal,n}(p_0,p_1,p_2)
	= \sum_{p_1,p_2 \in \Pcal_n}^{\ne} \ind{p_1 \in \BallPon{p_0}} \ind{p_2 \in \mathcal{T}_{\Pcal \Delta \Pcal_n}(p_0,p_1)},
\]
\TM{ Note you cannot really force/assume $p_0 = (0,y)$ is in your PPP $\Pcal_n$. (The summation over $\Pcal_n \setminus (0,y)$ suggests this is what you do.) For taking the expectation of sort of thing, Palm theory / Mecke are used typically. At this point though you can again still just say we add $p_0$ to the PPP. }\PvdH{I changed the text at the beginning of this section to mention this. Since we now consider the point process $\Pcal_n \cup p_0$ the above summation is over $\Pcal_n$.}
where the sums are over all distinct pairs $(p_1, p_2) \in \Pcal_n$.

@@

To further analyze this, let us introduce some notation. For any $p = (x,y) \in \Rcal_n$ we will define the left and right boundary functions as, respectively,
\begin{align}
	b_p^-(z) &= \begin{cases}
		2 \log\left(x-z\right) - y &\mbox{if }  -\frac{\pi}{2} e^{R_n/2} \le z \le x - e^{y/2}  \\
		2\log\left(\pi e^{R_n/2} + x - z\right) - y 
			&\mbox{if } x - e^{(y + R_n)/2} + \pi e^{R_n/2} \le z \le \frac{\pi}{2} e^{R_n/2}\\
		0 &\mbox{else}
	\end{cases}\\
	b_p^+(z) &= \begin{cases}
		2 \log\left(z-x\right) - y &\mbox{if } x + e^{y/2} \le z \le \frac{\pi}{2} e^{R_n/2} \\
		2\log\left(\pi e^{R_n/2} + z - x\right) - y 
			&\mbox{if } -\frac{\pi}{2} e^{R_n/2} \le z \le x + e^{(y + R_n)/2} - \pi e^{R_n/2}\\
		0 &\mbox{else}
	\end{cases}
\end{align}
Note that these functions describe the boundaries of the ball $\BallPon{p}$. In particular, $p^\prime = \BL{(x^\prime, y^\prime)} \in \BallPon{p}$ if and only if $y^\prime \ge \min\left\{b_p^-(x^\prime), b_p^+(x^\prime)\right\}$.

Since we have identified the left and right boundary of $\Rcal_n$ we can assume, without loss of generality that $x = 0$. Due to symmetry it is then enough to restrict the analysis to the case where $x^\prime > 0$. \TM{ does this not depend on $p$? }\PvdH{I do not think so, because we can always take $p = (0,y)$ due to the invariance in the $x$-direction. I have updated the text to better reflect this.} 
There are two important points in the box $\Rcal_n$. These are the intersection between the left boundary of $p^\prime$ and the right boundary of $p^\prime$, as it continues from the left side of the box, and the left boundary of $p$. We denote by $(x^\ast(p^\prime), y^\ast(p^\prime))$ the intersection between the left and right boundary of $p^\prime$ and by $(\hat{x}(p,p^\prime), \hat{y}(p,p^\prime))$ the intersection between the left boundary of $p$ and the right boundary of $p^\prime$, see Figure \ref{fig:comparing_triangles_diff_analysis}. After some simple algebra \TM{ I am not a big fan of writing "after some simple algebra" or similar. This is how errors can creep into your argument. Maybe we can just give the details? } we obtain the following expressions for the coordinates of these two points
\begin{align*}
	x^\ast(p^\prime) &= x^\prime - \frac{\pi}{2} e^{R_n/2}\\
	y^\ast(p^\prime) &= 2\log\left(\frac{\pi}{2}e^{R_n/2}\right) - y^\prime\\
	\hat{x}(p,p^\prime) &= \frac{x^\prime - \pi e^{R_n/2}}{1 + e^{(y^\prime - y)/2}} \\
	\hat{y}(p,p^\prime) &= 2 \log\left(\frac{\pi e^{R_n/2} - x^\prime}{e^{y/2} + e^{y^\prime/2}}\right)
\end{align*}
